DRAFT: Week 3: Degree and Centrality

### 2022-03-06 docs/3. SNA - Assignments/sna_Week3/DRAFT-week3-Assignment.Rmd
3.6.22

## Briefly describe the dataset:
1.   identify initial network format; 
2.   describe and identify the nodes 
     including: 
        *  how many nodes are in the dataset; 
3.   what constitutes a tie or edge
     including:
        *  how many ties, 
        *   whether ties are directed/undirected,
        *   whether ties are weighted/binary, 
        *   and how to interpret the value of the tie if any.
        *   Whether or not there are edge attributes 
            that might be used to:
            *   subset data, or 
            *   stack multiple networks.

@@

```{r Network Size}

  print(snEL)
  print(igEL)

  summary(snEL)
  summary(igEL)

```
plot.network(snEL)
igraph::plot.igraph(igEL)

###   IGRAPH:   Network Description
*igraph*
- Network Size
  - Number Vertex: 274
    - attr: name (v/c)
  - Number Edges: 672
    - attr: ingTyp (e/c)
- Bipartite: FALSE
- Directed: FALSE
- Weighted: FALSE

Names
- Vertex Names: view
- Edge Attribute Names: [1] "ingTyp"
```{r igraph: Network & Attribute Descripions, echo=TRUE}

@@

```{r iGraph: Undirected: network density, echo=TRUE}

    #get network density **with** loops
 igraph::graph.density(igEL)
 edge_density(igEL)
          ##**NOTE** provides an incorrect density calculation 
          ## as the default assumes that loops are not included but does not remove them

@@

```{r Statnet: Network Density, echo=TRUE}

            # get network density: with possible loops
   network.density(snEL)
          ##**WARNING: Network is multiplex - no general way to define density.**

 # get network density **with** loops: statnet
 gden(snEL, diag=F)

```{r statnet: Network Summary & Attribute info, echo=TRUE}
print(snEL)

@@

#      list the names of vertex attributes
  network::list.vertex.attributes(snEL)
#  Generate a list the names of edge attributes
  network::list.edge.attributes(snEL)

@@

plot(igEL_randoGraph,
      vertex.size = 5, 
      vertex.label = NA, 
      edge.curved = .1, 
      vertex.color = "tomato", 
      edge.arrow.size = .1, 
      edge.width = .5, 
     edge.color = "grey60")

@@

```
plot(igEL_triList)

@@

```
plot(view(
 sna::triad.census(
     snEL,
     mode="digraph")
))

@@

- Nodes: 
  - (v/c) Drink Names
  - (v/c) Drink Ingredients
- Links: 
  - (e/c)  Ingredient Type

@@

```
gplot(gd_snEL) in 
gplot(snEL_gdist)

@@

    K-Path Census
- Aggregate counts of paths/cycles by length (i.e., k).

@@

  The shortest path to describe the overall network structure 
```{r iGraph: Path: Network & Nodes, echo=TRUE}
library(igraph)

@@

 class(ig_histPath)
  class(ig_ShortPath)
  class(ig_AllShort)
```   
  Distance between two or more Nodes
```{r iGraph: Distances: Network & Nodes, echo=TRUE}

@@

### Statnet: Isolates & Component Structure
```{r Statnet: Isolates, echo=TRUE}
library(statnet)
library(sna)
              ## Isolates:

@@

##### Details: Component Size Distribution 
Components are maximal sets of mutually connected vertices; depending on the definition of “connected” one employs, one can arrive at several types of components. Those supported here are as follows (in increasing order of restrictiveness):

@@

 **Note:**

 the network density command ignores edge values at present.

 If loops are present or there are multiple edges between two nodes, then the results may not make sense.

```{r Statnet: Network Density, echo=TRUE}

            # get network density: with possible loops
   network.density(snEL)
          ##**WARNING: Network is multiplex - no general way to define density.**

 # get network density **with** loops: statnet

  gden(snEL, diag=F)

@@

```{r iGraph: edge density, echo=TRUE}
library(igraph)
    #get network density **with** loops
edge_density(igEL, loops = FALSE)

 class(igEL)
          ##**NOTE** provides an incorrect density calculation 
          ## as the default assumes that loops are not included but does not remove them

@@

## Vertex Degree
 **Vertex degree** is a description of an individual node; each vertex in a network may have a different node degree.

 *Node Degree:*
   node degree is one measure of node centrality or popularity; 
 Node degree cis a count of the relationships that a node is involved in or the count of the edges that are incident to a vertex

 **use the network degree command to:**
   *  compute network summary statistics (such as average network degree);
 *  graph the network degree distribution; or
 *  identify particular nodes (such as the five nodes with the most contacts).

```{r Statnet: Vertex Degree, echo=TRUE}
library(statnet)
library(sna)
      # statnet: Calculate average network degree
summary(sna::degree(snEL, 
             g=1, 
             nodes=NULL, 
             gmode="digraph", #directed;
             diag=FALSE, #excludes diagonals 
             tmaxdev=FALSE, 
             cmode="indegree", 
             rescale=FALSE, 
             ignore.eval=FALSE))
          # ignores loops or self-nominations 
summary(
  snEL_vd <- sna::degree(snEL, 
             g=1, 
             nodes= NULL, 
             gmode="graph", #undirected
             diag=TRUE, #includes diagonals 
             tmaxdev=F, 
             cmode="freeman", 
             rescale=F, #  scores rescaled to sum to 1.
             ignore.eval=FALSE))
plot(snEL_vd)
#**cmode=* degree centrality being computed. 
#*"indegree"
#*"outdegree"
#**DEFAULT*"freeman" refer to the indegree, outdegree, 
#*total (Freeman) degree measures, respectively. 
```

```{r iGraph: Undirected: Vertex Degree, echo=TRUE}
# igraph: Calculate average network degree
 # provides node name along with node degree

@@

## Network Degree Summary Statistics

```{r Statnet: , echo=TRUE}
library(statnet)
library(sna)   
   # Basic R: get summary statistics for node attributes

summary(snEL_Deg <- 
           sna::degree(snEL))
    plot(snEL_Deg)

      # provides range, mean, median and standard deviation of a vector.

@@

 6/274 vertices, named, from 1bfe0b4:
[1] (1-INCH-SQUARE) CHUNKS PINEAPPLE                     
[2] MINT LEAVES                                          
[3] LIME JUICE                                           
[4] DEMERARA SYRUP                                       
[5] PEACH LIQUEUR (SUCH AS MATHILDE, COMBIER, OR GIFFARD)
[6] BLENDED LIGHTLY AGED RUM

## Centrilization

```{r iGraph: Undirected: DEGREE CENTRALITY: dnkIng}
library(igraph)
library(tidyverse)

plot(centr_degree(igEL)$res)

@@

## Degree Centrality
We can break large social networks down into their constituent parts. These constituent parts are referred to as “motifs”. The most basic motif consists of two nodes and is called a dyad. Edges in a network signify the presence or absence of dyadic relations. It follows that a dyad in an undirected network can have two unique configurations: connected or disconnected; and three unique configurations in a directed network (mutual, assymetric, and null

```{r iGraph: Directed: GRAPH: Dyad Census, echo=TRUE}

@@

```{r iGraph: Directed: Degree Centrality}
# Degree centrality is calculated using the degree function in R. It returns how many edges each node has.
view(degree(igEL))

# plot(ingT_deg, 
#     vertex.label.cex = .6, 
#     vertex.label.color = "black", 
#     vertex.size = V(igEL)$degree, 
#     vertex.label.cex = .2) # sized by degree

```{r iGraph: Directed: Betweenness Centrality}

@@

Degree centrality only takes into account the number of edges for each node, but it leaves out information about ego’s alters
```{r iGraph: Directed: Eigenvector Centrality}

@@

 plot(ingTyp_Cent,
     vertex.label.cex = .6, 
     vertex.label.color = "black", 
     vertex.size = V(ingTyp_Cent)$page_rank/max(V(ingTyp_Cent)$page_rank) * 20)

>>>>

## Briefly describe the dataset:

1.  identify initial network format;

2.  describe and identify the nodes including:

    -   how many nodes are in the dataset;

3.  what constitutes a tie or edge including:

    -   how many ties,

    -   whether ties are directed/undirected,

    -   whether ties are weighted/binary,

    -   and how to interpret the value of the tie
        if any.

    -   Whether or not there are edge attributes
        that might be used to:

        -   subset data, or
        -   stack multiple networks.

@@

```{r Network Size & Info}

  print(snEL)
  summary(snEL)
  network::has.loops(snEL)
  network::is.directed(snEL)
  network::is.bipartite(snEL)
  network::is.hyper(snEL) # hypergraphic edges allowed
  network::is.multiplex(snEL) # multiplex edges allowed

  print_all(igEL)
  summary(igEL)
  igraph::is_directed(igEL)
  summary(
  igraph::is.loop(igEL))

plot.network(snEL) igraph::plot.igraph(igEL)

### IGRAPH: Network Description

*igraph* - Network Size - Number Vertex: 274 -
attr: name (v/c) - Number Edges: 672 - attr:
ingTyp (e/c) - Bipartite: FALSE - Directed: FALSE
- Weighted: FALSE

Names - Vertex Names: view - Edge Attribute Names:
[1] "ingTyp"

```{r igraph: Network & Attribute Descripions, echo=TRUE}

@@

@@

```{r statnet: Network Summary & Attribute info, echo=TRUE}

@@

  network::list.network.attributes(snEL)
#      list the names of vertex attributes
  network::list.vertex.attributes(snEL)
#  Generate a list the names of edge attributes
  network::list.edge.attributes(snEL)

@@

```{r iGraph: GRAPH: Dyad Census, echo=TRUE}

plot(
  igEL_randoGraph, 
  vertex.size = 5,
  vertex.label = NA, 
  edge.curved = .1, 
  vertex.color= "tomato", 
  edge.arrow.size = .1, edge.width = .5,
  edge.color = "grey60")

@@

plot(igEL_triList)

@@

plot(view( sna::triad.census( snEL,
mode="digraph") ))

@@

-   Nodes:

    -   (v/c) Drink Names
    -   (v/c) Drink Ingredients

-   Links:

    -   (e/c) Ingredient Type

@@

gplot(gd_snEL) in gplot(snEL_gdist)

@@

    K-Path Census

-   Aggregate counts of paths/cycles by length
    (i.e., k).

@@

The shortest path to describe the overall network
structure

```{r iGraph: Path: Network & Nodes, echo=TRUE}
library(igraph)

@@

 class(ig_histPath)
  class(ig_ShortPath)
  class(ig_AllShort)
```

Distance between two or more Nodes

```{r iGraph: Distances: Network & Nodes, echo=TRUE}

@@

### Statnet: Isolates & Component Structure

```{r Statnet: Isolates, echo=TRUE}
library(statnet)
library(sna)
              ## Isolates:

@@

##### Details: Component Size Distribution

Components are maximal sets of mutually connected
vertices; depending on the definition of
"connected" one employs, one can arrive at several
types of components. Those supported here are as
follows (in increasing order of restrictiveness):

@@

**Note:** the network density command ignores edge
values at present. If loops are present or there
are multiple edges between two nodes, then the
results may not make sense.

```{r Statnet: Network Density, echo=TRUE}
library(statnet)
library(sna)

  # computes network density
   network.density(snEL)

    network.density(
     snEL, 
     na.omit = T, # Exclude missing edges
     discount.bipartite = FALSE)

 # computes the density of  graphs and adjusts for the type of graph 
  gden(snEL)

  gden(
    snEL, 
       g=NULL, 
       diag=FALSE, # Set to TRUE if graph contains loops
       mode="graph", # digraph or graph
       ignore.eval=FALSE)

  network::is.bipartite(snEL)
  network::has.loops(snEL)

@@

```{r iGraph: network density / edge density, echo=TRUE}

library(igraph)

    #get network density **with** loops
 igraph::graph.density(
   igEL,
   loops = F)

 edge_density(igEL)
          ##**NOTE** provides an incorrect density calculation 
          ## as the default assumes that loops are not included but does not remove them

@@

## Network & Vertex Degree

### **use the network degree command to:**

-   compute network summary statistics (such as
    average network degree);
-   graph the network degree distribution; or
-   identify particular nodes (such as the five
    nodes with the most contacts). ##\# **Vertex
    degree** is a description of an individual
    node; each vertex in a network may have a
    different node degree.

###*Node Degree:* node degree is one measure of
node centrality or popularity; Node degree cis a
count of the relationships that a node is involved
in or the count of the edges that are incident to
a vertex

### Statnet: Network & Vertex Degree

#### Network Degree
```{r Statnet: Network Degree, echo=TRUE}
library(statnet)
library(sna)

 # statnet: Calculate average network degree
summary(  
  sna::degree(snEL))

summary(
  sna::degree(snEL, 
             g=1, 
             nodes=NULL, 
             gmode="digraph", #directed;
             diag=FALSE, # contains loops
             tmaxdev=FALSE, 
             cmode="indegree", 
             rescale=FALSE, 
             ignore.eval=FALSE)) # ignores loops or self-nominations 
summary(
  snEL_vd <- sna::degree(snEL, 
             g=1, 
             nodes= NULL, 
             gmode="graph", #undirected
             diag=TRUE, #includes diagonals 
             tmaxdev=F, 
             cmode="freeman", 
             rescale=F, #  scores rescaled to sum to 1.
             ignore.eval=FALSE))

plot(snEL_vd)
#**cmode=* degree centrality being computed. 
#*"indegree"
#*"outdegree"
#**DEFAULT*"freeman" refer to the indegree, outdegree, 
#*total (Freeman) degree measures, respectively. 
```

```{r Statnet: PLOT: Network Degree, echo=TRUE}
library(statnet)
library(sna)

 # statnet: Calculate average network degree

summary( snEL_netDeg <-
  sna::degree(snEL))

summary( snEL_vdDig  <-
 sna::degree(snEL, 
             g=1, 
             nodes=NULL, 
             gmode="digraph", #directed;
             diag=FALSE, # contains loops
             tmaxdev=FALSE, 
             cmode="indegree", 
             rescale=FALSE, 
             ignore.eval=FALSE)) # ignores loops or self-nominations 
summary(
  snEL_vdG <- sna::degree(snEL, 
             g=1, 
             nodes= NULL, 
             gmode="graph", #undirected
             diag=TRUE, #includes diagonals 
             tmaxdev=F, 
             cmode="freeman", 
             rescale=F, #  scores rescaled to sum to 1.
             ignore.eval=FALSE))

plot(snEL_netDeg)
plot(snEL_vdDig)
plot(snEL_vdG)

#**cmode=* degree centrality being computed. 
#*"indegree"
#*"outdegree"
#**DEFAULT*"freeman" refer to the indegree, outdegree, 
#*total (Freeman) degree measures, respectively. 
```

```{r Statnet: Vertex Degree (Directed Network), echo=TRUE}
library(statnet)
library(sna)

     ## create a dataset of vertex names and degree:
#####
snEL_nDeg <- data.frame(
  name=snEL%v%"vertex.names",
  degree=sna::degree(snEL))

snEL_nDegDig <-data.frame(
    name=snEL%v%"vertex.names",
    degree=sna::degree(snEL),
    indegree=sna::degree(snEL,
                         cmode="indegree"),
    outdegree=sna::degree(snEL,
                          cmode="outdegree"))

snEL_nInDegDig <-data.frame(
    name=snEL%v%"vertex.names",
    indegree=sna::degree(snEL,
                         cmode="indegree"))

snEL_nOutDegDig <-data.frame(
    name=snEL%v%"vertex.names",
    outdegree=sna::degree(snEL,
                          cmode="outdegree"))

@@

#### Network Degree Summary Statistics

```{r Statnet: Network Degree: Summary Stats, echo=TRUE}
library(statnet)
library(sna)

 # Basic R: get summary statistics for node attributes
 # provides range, mean, median and standard deviation of a vector.

  summary(snEL_nDeg)
  summary(snEL_nDegDig)
```

```{r Statnet: PLOT: Node Degree, echo=TRUE}
library(statnet)
library(sna)

@@

#### Network PLOT: Degree Distribution

hist() is a basic R command that draws a histogram
of vector of data

```{r Statnet: HIST: Network Degree Distribution}

library(statnet)
library(sna)

#create a histogram 
hist(
  snEL_nDegDig$degree, 
  main="Smuggler's Bay: Drink Names & Ingredients: Degree Distribution", 
  xlab="Nominations Received")

hist(
  snEL_nDegDig$indegree, 
  main="Smuggler's Bay: Drink Names & Ingredients: In-degree Distribution"
  , xlab="Nominations Received")

hist(
  snEL_nDegDig$outdegree, 
     main="Smuggler's Bay: Drink Names & Ingredients: Out-degree Distribution ", 
     xlab="Nominations made")

```{r Statnet: HIST: Node Degree Distribution, echo=TRUE}
library(statnet)
library(sna)

### AKU AKU   
snEL_degVNme <- data.frame(
   name=snEL%v%"Aku Aku",
    outdegree=sna::degree(snEL, cmode="outdegree"),
    indegree=sna::degree(snEL, cmode="indegree"))

hist(
  snEL_degVNme$outdegree, 
  main="Drink Name:  Aku Aku: Out Degree Distribution", 
  xlab="Nominations Received")

hist(
  snEL_degVNme$indegree, 
  main="Drink Name:  Aku Aku: Out Degree Distribution", 
  xlab="Nominations Received")

## "(1-INCH-SQUARE) CHUNKS PINEAPPLE"
snEL_degVNme <- data.frame(
   name=snEL%v%"(1-INCH-SQUARE) CHUNKS PINEAPPLE",
   indegree=sna::degree(snEL, cmode="indegree"),
   outdegree=sna::degree(snEL, cmode="outdegree"))

hist(
  snEL_degVNme$outdegree, 
  main="Drink Ingredient: (1-INCH-SQUARE) CHUNKS PINEAPPLE Out Degree Distribution", 
  xlab="Nominations Received")

hist(
  snEL_degVNme$indegree, 
  main="Drink Ingredient: Drink Ingredient: (1-INCH-SQUARE) CHUNKS PINEAPPLE in Degree Distribution", 
  xlab="Nominations Received")

## Mint Leaves 
snEL_degVNme <- data.frame(
   name=snEL%v%"MINT LEAVES",
   indegree=sna::degree(snEL, cmode="indegree"),
   outdegree=sna::degree(snEL, cmode="outdegree"))

hist(
  snEL_degVNme$outdegree, 
  main="Out Degree: Drink Ingredient: ", 
  xlab="Nominations Received")

hist(
  snEL_degVNme$indegree, 
  main="In Degree: Drink Ingredient: " , 
  xlab="Nominations Received")

## "DEMERARA SYRUP"
snEL_degVNme <- data.frame(
   name=snEL%v%"DEMERARA SYRUP",
   indegree=sna::degree(snEL, cmode="indegree"),
   outdegree=sna::degree(snEL, cmode="outdegree"))

hist(
  snEL_degVNme$outdegree, 
  main="Out Degree: Drink Ingredient: DEMERARA SYRUP", 
  xlab="Nominations Received")

hist(
  snEL_degVNme$indegree, 
  main="In Degree: Drink Ingredient:  DEMERARA SYRUP", 
  xlab="Nominations Received")

## Peach Liqueur
odeg_peachLqr <- data.frame(
   name=snEL%v%"PEACH LIQUEUR (SUCH AS MATHILDE, COMBIER, OR GIFFARD)",
   indegree=sna::degree(snEL, cmode="indegree"),
   outdegree=sna::degree(snEL, cmode="outdegree"))

hist(
  snEL_degVNme$outdegree, 
  main="Out Degree: Drink Ingredient: PEACH LIQUEUR", 
  xlab="Nominations Received")

hist(
  snEL_degVNme$indegree, 
  main="In Degree: Drink Ingredient: PEACH LIQUEUR" , 
  xlab="Nominations Received")

## " BLENDED LIGHTLY AGED RUM"
odeg_blaRum <- data.frame(
   name=snEL%v%"BLENDED LIGHTLY AGED RUM",
   indegree=sna::degree(snEL, cmode="indegree"),
   outdegree=sna::degree(snEL, cmode="outdegree"))

hist(
  snEL_degVNme$outdegree, 
  main="Out Degree: Drink Ingredient: BLENDED LIGHTLY AGED RUM ", 
  xlab="Nominations Received")

hist(
  snEL_degVNme$indegree, 
  main="In Degree: Drink Ingredient: BLENDED LIGHTLY AGED RUM " , 
  xlab="Nominations Received")

### iGraph: Vertex Degree

```{r iGraph: Vertex Degree, echo=TRUE}
# igraph: Calculate average network degree
 # provides node name along with node degree

## Network degree Centrilization
### Degree Centrality

We can break large social networks down into their
constituent parts. These constituent parts are
referred to as "motifs". The most basic motif
consists of two nodes and is called a dyad. Edges
in a network signify the presence or absence of
dyadic relations. It follows that a dyad in an
undirected network can have two unique
configurations: connected or disconnected; and
three unique configurations in a directed network
(mutual, assymetric, and null

### Statnet: Degree Centraility: In/Out

```{r Statnet: Network Degree centralization, eval=FALSE, include=FALSE}
library(statnet)
library(sna)

#get network centralization score:

# centralization(snEL, degree, cmode="indegree")

```{r Statmet: Netweork Centrality:  low & high degree nodes}

#sort to find 5 highest and lowest degree nodes in trade100

network::list.edge.attributes(snEL)

snEL_nodeHeadTail <-data.frame(
  name=snEL%v%"vertex.names",
  indegree=sna::degree(snEL, cmode="indegree"),
  outdegree=sna::degree(snEL, cmode="outdegree"))

snEL_nodeHeadTail%>%
  arrange(desc(indegree))%>%
  slice(1:10)

snEL_nodeHeadTail%>%
  arrange(indegree)%>%
  slice(1:10)

snEL_nodeHeadTail%>%
  arrange(desc(outdegree))%>%
  slice(1:10)

snEL_nodeHeadTail%>%
  arrange(outdegree)%>%
  slice(1:10)

view(snEL_nodeHeadTail)

### Igraph: Degree Centrality

```{r iGraph: DEGREE CENTRALITY}
library(igraph)
library(tidyverse)

#get network centralization score:

summary(
  centr_degree(igEL)$res)

  centr_degree(igEL)$centralization

summary(
  centr_degree(igEL)$centralizatio)

  centr_degree(igEL)$theoretical_max

#get network centralization score: in

  centr_degree(
    igEL, 
    loops = FALSE, 
    mode="in")$centralization

summary(
  centr_degree(
    igEL, 
    loops = FALSE, 
    mode="in")$centralization)

  centr_degree(
    igEL, 
    mode ="in")$theoretical_max

#get network centralization score: out

  centr_degree(
    igEL, 
    loops = FALSE, 
    mode="out")$centralization

summary(
  centr_degree(
    igEL, 
    loops = FALSE, 
    mode="out")$centralization)

  centr_degree(
    igEL,
    mode ="out")$theoretical_max

## get network centralization score: ALL

centr_degree(igEL,
  mode = c("all"),
  loops = TRUE,
  normalized = TRUE)$centralization

centr_degree(igEL,
  mode = c("all"),
  loops = TRUE,
  normalized = TRUE)$theoretical_max

```{r iGraph: PLOT: DEGREE CENTRALITY}
library(igraph)
library(tidyverse)

# network centralization score: 
plot(
  centr_degree(igEL)$res)

hist(
  centr_degree(igEL)$res
)

#get network centralization score: in

hist(
  centr_degree(igEL,mode = c("in"))$res)

  centr_degree(
    igEL, 
    loops = FALSE, 
    mode="in")$centralization

  centr_degree(igEL,
               mode ="in")$theoretical_max
#get network centralization score: out

  centr_degree(
    igEL, 
    loops = FALSE, 
    mode="out")$centralization

  centr_degree(
    igEL,
    mode ="out")$theoretical_max

  ##
summary(
  centr_degree(
    igEL, 
    loops = FALSE, 
    mode="out")$centralization)

@@

```{r iGraph: PLOT: Degree Centrality}
# Degree centrality is calculated using the degree function in R. It returns how many edges each node has.
view(degree(igEL))

plot(ingT_deg, 
     vertex.label.cex = .6, 
     vertex.label.color = "black", 
     vertex.size = V(igEL)$degree, 
     vertex.label.cex = .2) # sized by degree

@@

# WEEK4
```{r iGraph: Directed: Betweenness Centrality}

@@

Degree centrality only takes into account the
number of edges for each node, but it leaves out
information about ego's alters

```{r iGraph: Directed: Eigenvector Centrality}

@@

plot(ingTyp_Cent, vertex.label.cex = .6,
vertex.label.color = "black", vertex.size =
V(ingTyp_Cent)$page_rank/max(V(ingTyp_Cent)$page_rank)
\* 20)
